### 2017-08-19 cover/Cover-Letter.tex
25 pages version and cover letter

\documentclass[11pt]{article}

\usepackage[english]{babel}
\usepackage[latin1]{inputenc}
\usepackage{graphicx}
\usepackage{amssymb}
\usepackage{url}

%\usepackage{mathtools}
%\mathtoolsset{showonlyrefs}

\usepackage{color}
\usepackage{natbib}
% \usepackage[style=authoryear-comp,dashed=false]{biblatex}
% \bibliography{piecewise_lyapunov}

\textwidth 15cm
\setlength{\textheight}{1.1\textheight}
\newcommand\hi{\hspace*{\parindent}}
\newcommand\vi{\vspace{\baselineskip}}
\newcommand\lac{{\mbox{{\Huge\bf L}\hspace{-0.65em}
\raisebox{-1.2ex}{\Huge\bf A}\hspace{-1.1em}
\raisebox{-0.6ex}{\Huge\bf C~}}}}

\newcommand{\Real}{\mathbb{R}}

\pagestyle{empty}

\begin{document}

\hoffset -1.5cm \voffset .5cm

\sf

\vspace*{-2.5cm} \hspace*{-0.1cm} {\small
{\mbox{\begin{minipage}{1.5cm}
\centerline{\includegraphics[width=1.8cm]{oxford.jpg}}
\end{minipage}}} \hspace*{0.1cm}
\begin{minipage}{11.0cm}{\sf \large \textbf{UNIVERSITY OF OXFORD}}\\
{\sc Department of Computer Science}
\end{minipage}
} \vspace*{0mm}

\hspace*{-.7cm} {\rule[-1ex]{15cm}{0.03cm}}

\vi \vi
\begin{flushright}
\begin{minipage}{7.0cm}\small
{\bf Please Reply to:}\\
Dr. Cordeiro\\
University of Oxford\\
Department of Computer Science\\
Wolfson Building, Parks Rd, Oxford OX1 3QD\\
{\em e-mail: lucas.cordeiro@cs.ox.ac.uk}
\end{minipage}
\end{flushright}

\vi
\hspace*{\fill}{\small Oxford, $19^{th}$ August, 2017.}
\vi

\begin{flushleft}
    Tei-Wei Kuo\\
    Editor-in-Chief\\
    ACM Transactions on Cyber-Physical Systems (TCPS)
    \end{flushleft}
\vi

\indent Dear Editor-in-Chief,
\vi

Please find enclosed our manuscript entitled as ``{\em Automated Program Verification and Synthesis of Embedded \& Cyber-physical Systems: Advances and Challenges},'' by Lucas Cordeiro, Eddie Batista de Lima Filho, and Iury Bessa, which we would like to submit for publication in the journal of {\bf ACM Transactions on Cyber-Physical Systems (TCPS)}.

\vi

The present work is a survey paper that discusses challenges, problems, and recent advances to ensure correctness and timeliness regarding ECPSs. Reliability issues, in the development of ECPSs, are then considered, as a prominent verification and synthesis application for achieving a correct-by-construction design.

\vi

The  authors  claim  that  none  of  the  material  in  the mentioned  paper  has  been  published  or  is  under  consideration for publication elsewhere.

\vi

Thank you for your time.

\vi\vi

\indent
~~~~~~~~~~~~~~~~~~~~~~~~~~~Sincerely,\\

\begin{quote}
\begin{quote}
\begin{flushright}

%\vfill

Lucas C. Cordeiro~~~~~~~
\end{flushright}
\end{quote}
\end{quote}

\end{document}